---
title: "Large Hurricane Analysis - Precipitation"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{top_15}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

## Program Background

-   Script name: Large Hurricane Analysis - Precipitation

-   Project name: ADRD and Hurricane Exposure

-   Program description:

    Amount of precipitation was calculated for every day specified within the HURDAT2 plus an additional 2 days after. nClimGrid raster data was used to calculate an areal mean precipitation for all tracts/zctas.

-   Data:

    -   HURDAT2 - to get storm dates
    -   [nClimGrid-Daily](https://www.ncei.noaa.gov/products/land-based-station/nclimgrid-daily)

-   Restrictions:

-   Outline of program/analysis:

    -   Specify all storms
    -   Download and process hurdat2 tracks
    -   Download nClimGrid data for duration of storm plus 2 days after
    -   Calculate areal mean precipitation

-   Author: Ryan Zomorrodi

-   Revisions

-   Input files

    -   All files are downloaded using R functions
    -   Cache is stored in `raw-data\nClimGrid`

-   Output files

    -   `tract/{storm id}/daily_prcp.csv`

    -   `zcta/{storm id}/daily_prcp.csv`

        +-----------------+-----------------------------------------------------------------------------------------------------------+
        | Column          | Description                                                                                               |
        +=================+===========================================================================================================+
        | `geoid`         | 2010 geographic identifier                                                                                |
        +-----------------+-----------------------------------------------------------------------------------------------------------+
        | `date`          | Date in UTC                                                                                               |
        +-----------------+-----------------------------------------------------------------------------------------------------------+
        | `mean_prcp`     | Mean areal precipitation (mm) for the specified `geoid` and `tract`                                       |
        |                 |                                                                                                           |
        |                 | Note: there may be NAs for a small number of coastline tract which were not entirely covered by nClimGrid |
        +-----------------+-----------------------------------------------------------------------------------------------------------+

```{r setup}
library(exposur)
library(tidyverse)
library(tigris)
library(sf)
```

The following hurricanes were selected.

```{r}
# manually selected states of interest for each storm
hur_info <- tribble(
  ~ hurricane, ~ states,
  "KATRINA-2005",  c("LA", "MS", "AL", "FL", "TN"),
  "HARVEY-2017",   c("TX", "LA"),
  "SANDY-2012",    c("NJ", "DE", "MD", "NY", "CT", "RI", "VA", "MA", "PA"),
  "IRMA-2017",     c("FL", "GA", "SC", "AL"),
  "IKE-2008",      c("TX", "LA"),
  "IVAN-2004",     c("AL", "MS", "FL", "GA", "TN", "NC", "SC"),
  "RITA-2005",     c("TX", "LA"),
  "FLORENCE-2018", c("NC", "SC", "GA", "VA"),
  "MICHAEL-2018",  c("FL", "GA", "AL"),
  "WILMA-2005",    c("FL"),
  "CHARLEY-2004",  c("FL", "GA", "SC", "NC"),
  "IRENE-2011",    c("SC", "NC", "VA", "MD", "DE", "NJ", "PA", "NY", "CT"),
  "FRANCES-2004",  c("FL", "GA", "SC", "NC", "AL", "TN", "VA"),
  "ALLISON-2001",  c("TX", "LA", "MS", "AL", "FL"),
  "MATTHEW-2016",  c("FL", "GA", "SC", "NC", "VA")
)
# get fips codes for each state
fips_codes <- distinct(tigris::fips_codes, state, state_code)
hur_info$fips <- pmap(hur_info, function(states, ...) {
  fips_codes$state_code[match(states, fips_codes$state)]
})
```

## Get hurdat2 info for each storm

```{r}
exposur:::download_newest_hurdat2(path = "raw-data/hurdat2")

hur_info$hurdat2 <- pmap(hur_info, function(hurricane, ...){
  hurdat  <- get_hurdat2(
    storm_id = hurricane,
    path = "raw-data/hurdat2"
  )
  
  list(
    start_date = min(hurdat$datetime),
    end_date = max(hurdat$datetime),
    hurdat = hurdat
  )
})
hur_info <- unnest_wider(hur_info, hurdat2)
```

## Get states

```{r, warning = FALSE}
hur_info$tracts_sf <- pmap(hur_info, function(start_date, fips, ...) {
  tracts_list <- map(fips, 
    ~ tracts(state = .x, year = 2010, cb = TRUE)
  )
  bind_rows(tracts_list) %>%
    mutate(GEOID = str_sub(GEO_ID, -11)) %>% # get tract geoids
    filter(!st_is_empty(.)) %>% # remove tracts with empty geographies
    st_transform(4326) %>% # transform to WGS 84
    rename(geoid = GEOID)
})

hur_info$zctas_sf <- pmap(hur_info, function(start_date, fips, ...) {
  zctas_list <- map(fips, 
    ~ zctas(state = .x, year = 2010, cb = FALSE)
  )
  bind_rows(zctas_list) %>%
    filter(!st_is_empty(.)) %>% # remove zctas with empty geographies
    st_transform(4326) %>%
    rename(geoid = GEOID10)
})
```

## Get nClimGrid

```{r}
hur_info$nClimGrid <- pmap(hur_info, function(start_date, end_date, ...) {
  get_nClimGrid(
    interval = interval(start_date, end_date + days(2)),
    path = "raw-data/nClimGrid"
  )["prcp"]
})
```

## Calculate daily precipitation

```{r}
hur_info$tract_prcp_daily_avg <- pmap(hur_info, function(nClimGrid, tracts_sf, ...) {
  calc_daily_avgs(
    nClimGrid = nClimGrid,
    geography = tracts_sf,
    geoid = "geoid"
  )
})

pwalk(hur_info, function(tract_prcp_daily_avg, hurricane, ...) {
  write_csv(tract_prcp_daily_avg, str_c("tract/", hurricane, "/daily_prcp.csv"))
})

hur_info$zcta_prcp_daily_avg <- pmap(hur_info, function(nClimGrid, zctas_sf, ...) {
  calc_daily_avgs(
    nClimGrid = nClimGrid,
    geography = zctas_sf,
    geoid = "geoid"
  )
})

pwalk(hur_info, function(zcta_prcp_daily_avg, hurricane, ...) {
  write_csv(zcta_prcp_daily_avg, str_c("zcta/", hurricane, "/daily_prcp.csv"))
})
```

End Time

```{r}
end_time <- Sys.time()
cat("End time: ", end_time, "\n")
cat("Duration: ", start_time - end_time, "\n")
```