---
title: "Large Hurricane Analysis - Distance and Wind"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{top_15}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Program Background

-   Script name: Large Hurricane Analysis - Distance and Wind

-   Project name: ADRD and Hurricane Exposure

-   Program description:

    Maximum wind was calculated for each tract using the `stormwindmodel` published by Anderson at the geocentroid for each tract/zcta.

-   Data:

    -   HURDAT2

-   Restrictions:

-   Outline of program/analysis:

    -   Specify all storms
    -   Download and process hurdat2 tracks
    -   Calculate geo-centroid for each tract/zcta
    -   Utilize stormwindmodel to produce wind exposure estimates
    -   Output imputed track geojsons for visualization

-   Author: Ryan Zomorrodi

-   Revisions

-   Input files

    -   All files are downloaded using R functions
    -   Cache is stored in `raw-data\hurdat2`

-   Output files

    -   `tract/{storm_id}/distance_wind.csv`

    -   `zcta/{storm_id}/distance_wind.csv`

        | Column | Description |
        |------------------------------------|------------------------------------|
        | geoid | 2010 geographic identifier |
        | closest_datetime | Date and time in UTC that storm is closest to the imputed storm track |
        | closest_dist | Closest distance (km) to imputed central track |
        | highwind_datetime | Date and time in UTC of maximum wind speed |
        | vmax_sust | Maximum 10-m 1-min sustained wind for the tropical cyclone (m/s) |
        | vmax_gust | Max 10-m 1-min gust wind experienced at grid point (m/s) |
        | sust_dur | Duration of time a certain sustained wind was experienced at grid point (minutes) |
        | gust_dur | Duration of time a certain gust wind was experienced at grid point (minutes) |

    -   `tract/{storm_id}/track_line.geojson`

    -   `zcta/{storm_id}/track_line.geojson`

    -   `tract/{storm_id}/track_points.geojson`

    -   `zcta/{storm_id}/track_points.geojson`

Start Time

```{r, eval = FALSE}
start_time <- Sys.time()
cat(start_time, "\n")
```

```{r setup, eval = FALSE}
library(exposur)
library(tidyverse)
library(tigris)
library(sf)
```

The following hurricanes were selected.

```{r, eval = FALSE}
hur_info <- readRDS( "raw-data/RF1_Hurricanes.rds")
# get fips codes for each state
fips_codes <- distinct(tigris::fips_codes, state, state_code)
hur_info$fips <- pmap(hur_info, function(states, ...) {
  fips_codes$state_code[match(states, fips_codes$state)]
})
```

## Get hurdat2 info for each storm

```{r, eval = FALSE}
exposur:::download_newest_hurdat2(path = "raw-data/hurdat2")

hur_info$hurdat2 <- pmap(hur_info, function(hurricane, ...){
  hurdat  <- get_hurdat2(
    storm_id = hurricane,
    path = "raw-data/hurdat2"
  )
  
  list(
    start_date = min(hurdat$datetime),
    end_date = max(hurdat$datetime),
    hurdat = hurdat
  )
})
hur_info <- unnest_wider(hur_info, hurdat2)
```

## Get states

```{r, warning = FALSE, eval = FALSE}
hur_info$tracts_sf <- pmap(hur_info, function(start_date, fips, ...) {
  tracts_list <- map(fips, 
    ~ tracts(state = .x, year = 2010, cb = TRUE)
  )
  bind_rows(tracts_list) %>%
    mutate(GEOID = str_sub(GEO_ID, -11)) %>% # get tract geoids
    filter(!st_is_empty(.)) %>% # remove tracts with empty geographies
    st_transform(4326) %>% # transform to WGS 84
    rename(geoid = GEOID) %>%
    st_centroid() # get geo-centroid of each tract
})

hur_info$zctas_sf <- pmap(hur_info, function(start_date, fips, ...) {
  zctas_list <- map(fips, 
    ~ zctas(state = .x, year = 2010, cb = FALSE)
  )
  bind_rows(zctas_list) %>%
    filter(!st_is_empty(.)) %>% # remove zctas with empty geographies
    st_transform(4326) %>%
    rename(geoid = GEOID10) %>%
    st_centroid() # get geo-centroid of each zctas
})
```

## Calculate interpolated track

```{r, eval = FALSE}
hur_info$track <- pmap(hur_info, function(hurdat, ...) {
  calc_interpolated_track(hurdat)
})

pwalk(hur_info, function(track, hurricane, ...) {
  st_write(track, str_c("tract/", hurricane, "/track_points.geojson"))
  
  track %>%
    st_combine() %>%
    st_cast("LINESTRING") %>%
    st_write(str_c("tract/", hurricane, "/track_line.geojson"))
})

pwalk(hur_info, function(track, hurricane, ...) {
  st_write(track, str_c("zcta/", hurricane, "/track_points.geojson"))
  
  track %>%
    st_combine() %>%
    st_cast("LINESTRING") %>%
    st_write(str_c("zcta/", hurricane, "/track_line.geojson"))
})
```

## Calculate closest distance

```{r, eval = FALSE}
hur_info$tract_closest_dist <- pmap(hur_info, function(track, tracts_sf, ...) {
  calc_closest_dist(track, tracts_sf, geoid = "geoid")
})

hur_info$zcta_closest_dist <- pmap(hur_info, function(track, zctas_sf, ...) {
  calc_closest_dist(track, zctas_sf, geoid = "geoid")
})
```

## Calculate wind speed

```{r, eval = FALSE}
hur_info$tract_anderson_wind <- pmap(hur_info, function(hurdat, tracts_sf, ...) {
  calc_anderson_wind(hurdat, tracts_sf)
})

hur_info$zcta_anderson_wind <- pmap(hur_info, function(hurdat, zctas_sf, ...) {
  calc_anderson_wind(hurdat, zctas_sf)
})
```

## Join output and write to file

```{r, eval = FALSE}
pwalk(hur_info, function(tract_closest_dist, tract_anderson_wind, hurricane, ...) {
  full_join(tract_closest_dist, tract_anderson_wind, join_by(geoid)) %>%
    relocate(closest_date, .before = closest_dist) %>%
    rename(highwind_datetime = datetime, closest_datetime = closest_date) %>%
    write_csv(str_c("tract/", hurricane, "/distance_wind.csv"))
})
 
pwalk(hur_info, function(zcta_closest_dist, zcta_anderson_wind, hurricane, ...) {
  full_join(zcta_closest_dist, zcta_anderson_wind, join_by(geoid)) %>%
    relocate(closest_date, .before = closest_dist) %>%
    rename(highwind_datetime = datetime, closest_datetime = closest_date) %>%
    write_csv(str_c("zcta/", hurricane, "/distance_wind.csv"))
})
```

End Time

```{r, eval = FALSE}
end_time <- Sys.time()
cat("End time: ", end_time, "\n")
cat("Duration: ", start_time - end_time, "\n")
```

## Visualizations

```{r, include = FALSE}
library(tidyverse)
library(tigris)
library(sf)

# repeat from above
hur_info <- readRDS( "../raw-data/RF1_Hurricanes.rds")
fips_codes <- distinct(tigris::fips_codes, state, state_code)
hur_info$fips <- pmap(hur_info, function(states, ...) {
  fips_codes$state_code[match(states, fips_codes$state)]
})

hur_info$tracts_sf <- pmap(hur_info, function(fips, ...) {
  tracts_list <- map(fips, 
    ~ tracts(state = .x, year = 2010, cb = TRUE)
  )
  bind_rows(tracts_list) %>%
    mutate(GEOID = str_sub(GEO_ID, -11)) %>%
    filter(!st_is_empty(.))
})

hur_info$states_sf <- pmap(hur_info, function(fips, ...) {
  states(year = 2010, cb = TRUE) %>%
    filter(STATE %in% fips)
})

hur_info$track_line <- pmap(hur_info, function(hurricane, ...) {
  st_read(str_c("../tract/", hurricane, "/track_line.geojson"))
})

hur_info$tract_anderson_wind <- pmap(hur_info, function(hurricane, ...) {
  read_csv(
      str_c("../tract/", hurricane, "/distance_wind.csv"),
      col_types = cols(geoid = col_character())
  )
})
```

```{r, fig.width = 10, fig.height = 10, echo = FALSE}
library(patchwork)
library(classInt)

hur_info$wind_plots <- pmap(hur_info, function(hurricane, tracts_sf, tract_anderson_wind, states_sf, track_line,  ...) {
  wind_sf <- left_join(
    tracts_sf, 
    tract_anderson_wind, 
    by = join_by(GEOID == geoid)
  ) %>%
    mutate(vmax_sust = vmax_sust * 3.6)
  
  ggplot(wind_sf) +
    geom_sf(aes(fill = vmax_sust), color = NA) +
    geom_sf(data = states_sf, fill = NA) +
    geom_sf(
      data = st_intersection(
        st_transform(track_line, st_crs(states_sf)), 
        st_as_sfc(st_bbox(states_sf))
      )
    ) +
    scale_fill_distiller(palette = "Greens", na.value= "grey", direction = 1) +
    coord_sf(crs = "ESRI:102003") +
    theme_bw() +
    theme(
      axis.text.x = element_blank(),
      axis.text.y = element_blank(),
      axis.ticks = element_blank()
    ) +
    labs(title = hurricane, fill = "")
})

wrap_plots(
  hur_info$wind_plots,
  ncol = 3
) +
  plot_annotation(
    title = "Estimated Maximum Wind Velocity (km/h)",
    subtitle = "stormwindmodel estimates at 2010 tract geocentroid",
    theme = theme(plot.title = element_text(size = 16))
  )
```

```{r, echo = FALSE}
library(gt)

hur_info$dist_table <- pmap(hur_info, function(hurricane, tract_anderson_wind, ...) {
  
  tract_anderson_wind %>%
    mutate(vmax_sust = vmax_sust * 3.6) %>%
    summarize(
      Storm = hurricane,
      Mean = mean(vmax_sust, na.rm = TRUE),
      SD = sd(vmax_sust, na.rm = TRUE),
      Min = min(vmax_sust, na.rm = TRUE),
      `25% Quantile` = quantile(vmax_sust, probs = 0.25, na.rm = TRUE),
      Median = quantile(vmax_sust, probs = 0.5, na.rm = TRUE),
      `75% Quantile` = quantile(vmax_sust, probs = 0.75, na.rm = TRUE),
      Max = max(vmax_sust, na.rm = TRUE),
      `Proportion 34+ knots` = mean(vmax_sust >= 62.968, na.rm = TRUE),
      Distribution = list(vmax_sust)
    ) %>%
    mutate(across(where(is.numeric), ~ round(.x, 3)))
})
    
bind_rows(hur_info$dist_table) %>%
  gt() %>%
  gtExtras::gt_plt_dist(Distribution) %>%
  tab_header(
    title = "Estimated Maximum Wind Velocity (km/h)",
    subtitle = "stormwindmodel estimates at 2010 tract geocentroid",
  )
```

```{r, fig.width = 10, fig.height = 10, echo = FALSE}
library(patchwork)
library(classInt)

hur_info$dist_plots <- pmap(hur_info, function(hurricane, tracts_sf, tract_anderson_wind, states_sf, track_line,  ...) {
  dist_sf <- left_join(
    tracts_sf, 
    tract_anderson_wind, 
    by = join_by(GEOID == geoid)
  )
  
  ggplot(dist_sf) +
    geom_sf(aes(fill = closest_dist), color = NA) +
    geom_sf(data = states_sf, fill = NA) +
    geom_sf(
      data = st_intersection(
        st_transform(track_line, st_crs(states_sf)), 
        st_as_sfc(st_bbox(states_sf))
      )
    ) +
    scale_fill_distiller(palette = "Purples", na.value= "grey", direction = -1) +
    coord_sf(crs = "ESRI:102003") +
    theme_bw() +
    theme(
      axis.text.x = element_blank(),
      axis.text.y = element_blank(),
      axis.ticks = element_blank()
    ) +
    labs(title = hurricane, fill = "")
})

wrap_plots(
  hur_info$dist_plots,
  ncol = 3
) +
   plot_annotation(
    title = "Estimated Closest Approach (km)",
    subtitle = "Smallest distance from 2010 tract geocentroid to points on 15 min imputed track",
    theme = theme(plot.title = element_text(size = 16))
  )
```

```{r, echo = FALSE}
library(gt)

hur_info$dist_table <- pmap(hur_info, function(hurricane, tract_anderson_wind, ...) {
  summarize(
    tract_anderson_wind,
    Storm = hurricane,
    Mean = mean(closest_dist, na.rm = TRUE),
    SD = sd(closest_dist, na.rm = TRUE),
    Min = min(closest_dist, na.rm = TRUE),
    `25% Quantile` = quantile(closest_dist, probs = 0.25, na.rm = TRUE),
    Median = quantile(closest_dist, probs = 0.5, na.rm = TRUE),
    `75% Quantile` = quantile(closest_dist, probs = 0.75, na.rm = TRUE),
    Max = max(closest_dist, na.rm = TRUE),
    Distribution = list(closest_dist)
  ) %>%
    mutate(across(where(is.numeric), ~ round(.x, 3)))
})
    
bind_rows(hur_info$dist_table) %>%
  gt() %>%
  gtExtras::gt_plt_dist(Distribution) %>%
  tab_header(
    title = "Estimated Closest Approach (km)",
    subtitle = "Smallest distance from 2010 tract geocentroid to points on 15 min imputed track"
  )
```