---
title: "Large Hurricane Analysis - Outage"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{top_15}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Program Background

-   Script name: Large Hurricane Analysis - Outage

-   Project name: ADRD and Hurricane Exposure

-   Program description:

    Outage measure calculation for 15 previously selected large hurricanes. Only hurricanes in 2014 and after had available outage data.

    Number of customers without electricity is collected by EAGLE-I. County-level electric customer estimates are generating by adding the census business survey and ACS5 population estimates for a county and normalizing to state-wide energy customer estimates from the Energy Information Administration.

-   Data:

    -   [EAGLE-I outage data](https://figshare.com/ndownloader/articles/24237376/versions/2)

    -   [Census Business Survey](https://www2.census.gov/programs-surveys/susb/tables/)

    -   ACS 5 - through `tidycensus` package

    -   [EIA State Customer Estimates](https://www.eia.gov/electricity/data/browser/data/index.php?method=getMapData&geosetId=ELEC.CUSTOMERS.ALL.A&freq=A&topic=56)

    -   HURDAT2 - to get storm dates

    -   [2010 ZCTA to County Relationships](https://www2.census.gov/geo/docs/maps-data/data/rel/zcta_county_rel_10.txt)

-   Restrictions:

-   Outline of program/analysis:

    -   Specify all storms

    -   Download and process hurdat2 tracks

    -   Extract outage data for dates specified by the HURDAT2 plus 2 days after

    -   Calculate outage estimates for the maximum percentage of customers without electricity and the average duration of outages

    -   Crosswalk county estimates to ZCTA using 2010 ZCTA to County Relationship file

-   Author: Ryan Zomorrodi

-   Revisions

-   Input files

    -   All files are downloaded using R functions
    -   Cache is stored in `raw-data\eaglei`

-   Output files

    -   `tract/{storm id}/outage.csv`

    -   `zcta/{storm id}/outage.csv`

        | Column | Description |
        |------------------------------------|------------------------------------|
        | `geoid` | 2010 geographic identifier |
        | `max_pct_out` | Maximum percentage of customers out of electricity at any point |
        | `avg_customer_hrs_out` | Average number of hours each customer was out of electricity for |

Start Time

```{r, eval = FALSE}
start_time <- Sys.time()
cat(start_time, "\n")
```

```{r setup, eval = FALSE}
library(exposur)
library(tidyverse)
library(tigris)
library(sf)
```

The following hurricanes were selected.

```{r, eval = FALSE}
hur_info <- readRDS( "raw-data/RF1_Hurricanes.rds")
# get fips codes for each state
fips_codes <- distinct(tigris::fips_codes, state, state_code)
hur_info$fips <- pmap(hur_info, function(states, ...) {
  fips_codes$state_code[match(states, fips_codes$state)]
})
```

## Get hurdat2 info for each storm

```{r, eval = FALSE}
exposur:::download_newest_hurdat2(path = "raw-data/hurdat2")

hur_info$hurdat2 <- pmap(hur_info, function(hurricane, ...){
  hurdat  <- get_hurdat2(
    storm_id = hurricane,
    path = "raw-data/hurdat2"
  )
  
  list(
    start_date = min(hurdat$datetime),
    end_date = max(hurdat$datetime),
    hurdat = hurdat
  )
})
hur_info <- unnest_wider(hur_info, hurdat2)
```

## Get outage data

```{r, eval = FALSE}
hur_info$eaglei <- pmap(hur_info, function(start_date, end_date, ...) {
  if (year(start_date) < 2014) {
    return(list(NULL))
  }

  get_eaglei(
    interval = interval(start_date, end_date + days(2)),
    path = "raw-data/eaglei"
  )
})
```

## Calculate outage

```{r, eval = FALSE}
hur_info$outage <- pmap(hur_info, function(eaglei, fips, start_date, ...) {
  if (year(start_date) < 2014) {
    return(list(NULL))
  }

  calc_outage(eaglei, fips)
})

pwalk(hur_info, function(outage, hurricane, start_date, ...) {
  if (year(start_date) >= 2014) {
    write_csv(outage, str_c("tract/", hurricane, "/outage.csv"))
  }
})
```

## Crosswalk to zctas

```{r, eval = FALSE}
zcta_to_county <- read_csv("https://www2.census.gov/geo/docs/maps-data/data/rel/zcta_county_rel_10.txt") %>%
  mutate(geoid_z = str_c(STATE, ZCTA5)) %>%
  mutate(geoid_c = str_c(STATE, COUNTY))

outages <- Sys.glob("tract/*/outage.csv") %>%
  map(function(outage_path) {
    outage <- read_csv(outage_path, col_types = cols(geoid = col_character()))
    
    outage %>%
      left_join(zcta_to_county, join_by(geoid == geoid_c)) %>%
      group_by(geoid_z) %>%
      summarise(
        max_pct_out_xwalk = sum(ZPOPPCT * max_pct_out) / 100,
        avg_customer_hrs_out_xwalk = sum(ZPOPPCT * avg_customer_hrs_out) / 100
      )
    
    outage %>%
      write_csv(str_replace(outage_path, "tract", "zcta"))
  })
```

End Time

```{r, eval = FALSE}
end_time <- Sys.time()
cat("End time: ", end_time, "\n")
cat("Duration: ", start_time - end_time, "\n")
```

## Visualizations

```{r, include = FALSE}
library(tidyverse)
library(tigris)
library(sf)

# repeat from above
hur_info <- readRDS( "../raw-data/RF1_Hurricanes.rds") %>%
  filter(year >= 2014)
fips_codes <- distinct(tigris::fips_codes, state, state_code)
hur_info$fips <- pmap(hur_info, function(states, ...) {
  fips_codes$state_code[match(states, fips_codes$state)]
})

hur_info$outage <- pmap(hur_info, function(hurricane, ...) {
  read_csv(
    str_c("../tract/", hurricane, "/outage.csv"),
    col_types = cols(geoid = col_character())
  )
})

hur_info$counties_sf <- pmap(hur_info, function(fips, ...) {
  counites_list <- map(fips, 
    ~ counties(state = .x, year = 2010, cb = TRUE)
  )
  bind_rows(counites_list) %>%
    mutate(GEOID = str_sub(GEO_ID, -5)) %>%
    filter(!st_is_empty(.))
})

hur_info$states_sf <- pmap(hur_info, function(fips, ...) {
  states(year = 2010, cb = TRUE) %>%
    filter(STATE %in% fips)
})

hur_info$track_line <- pmap(hur_info, function(hurricane, ...) {
  st_read(str_c("../tract/", hurricane, "/track_line.geojson"))
})
```

```{r, fig.width = 10, fig.height = 6, echo = FALSE, warning = FALSE}
library(patchwork)
library(classInt)

hur_info$out_plots <- pmap(hur_info, function(hurricane, counties_sf, outage, states_sf, track_line,  ...) {
  out_sf <- left_join(
    counties_sf, 
    outage, 
    by = join_by(GEOID == geoid)
  ) %>%
    mutate(max_pct_out = max_pct_out * 100)
  
  ggplot(out_sf) +
    geom_sf(aes(fill = max_pct_out), color = NA) +
    geom_sf(data = states_sf, fill = NA) +
    geom_sf(
      data = st_intersection(
        st_transform(track_line, st_crs(states_sf)), 
        st_as_sfc(st_bbox(states_sf))
      )
    ) +
    scale_fill_distiller(palette = "Oranges", na.value= "grey", direction = 1) +
    coord_sf(crs = "ESRI:102003") +
    theme_bw() +
    theme(
      axis.text.x = element_blank(),
      axis.text.y = element_blank(),
      axis.ticks = element_blank()
    ) +
    labs(title = hurricane, fill = "")
})

wrap_plots(
  hur_info$out_plots,
  ncol = 3
) +
   plot_annotation(
    title = "Estimated Maximum Percentage of Customers without Electricity",
    theme = theme(plot.title = element_text(size = 16))
  )
```

```{r, echo = FALSE}
library(gt)

hur_info$out_table <- pmap(hur_info, function(hurricane, outage, ...) {
  outage %>%
    mutate(max_pct_out = max_pct_out * 100) %>%
    summarize(
      Storm = hurricane,
      Mean = mean(max_pct_out, na.rm = TRUE),
      SD = sd(max_pct_out, na.rm = TRUE),
      Min = min(max_pct_out, na.rm = TRUE),
      `25% Quantile` = quantile(max_pct_out, probs = 0.25, na.rm = TRUE),
      Median = quantile(max_pct_out, probs = 0.5, na.rm = TRUE),
      `75% Quantile` = quantile(max_pct_out, probs = 0.75, na.rm = TRUE),
      Max = max(max_pct_out, na.rm = TRUE),
      Distribution = list(max_pct_out)
    ) %>%
    mutate(across(where(is.numeric), ~ round(.x, 3)))
})
    
bind_rows(hur_info$out_table) %>%
  gt() %>%
  gtExtras::gt_plt_dist(Distribution) %>%
  tab_header(
    title = "Estimated Maximum Percentage of Customers without Electricity",
  )
```


```{r, fig.width = 10, fig.height = 6, echo = FALSE, warning = FALSE}
library(patchwork)

hur_info$out_plots <- pmap(hur_info, function(hurricane, counties_sf, outage, states_sf, track_line,  ...) {
  out_sf <- left_join(
    counties_sf, 
    outage, 
    by = join_by(GEOID == geoid)
  )
  
  ggplot(out_sf) +
    geom_sf(aes(fill = avg_customer_hrs_out), color = NA) +
    geom_sf(data = states_sf, fill = NA) +
    geom_sf(
      data = st_intersection(
        st_transform(track_line, st_crs(states_sf)), 
        st_as_sfc(st_bbox(states_sf))
      )
    ) +
    scale_fill_distiller(palette = "Reds", na.value= "grey", direction = 1) +
    coord_sf(crs = "ESRI:102003") +
    theme_bw() +
    theme(
      axis.text.x = element_blank(),
      axis.text.y = element_blank(),
      axis.ticks = element_blank()
    ) +
    labs(title = hurricane, fill = "")
})

wrap_plots(
  hur_info$out_plots,
  ncol = 3
) +
   plot_annotation(
    title = "Estimated Average Customer Hours without Electricity",
    theme = theme(plot.title = element_text(size = 16))
  )
```

```{r, echo = FALSE}
library(gt)

hur_info$out_table <- pmap(hur_info, function(hurricane, outage, ...) {
  outage %>%
    summarize(
      Storm = hurricane,
      Mean = mean(avg_customer_hrs_out, na.rm = TRUE),
      SD = sd(avg_customer_hrs_out, na.rm = TRUE),
      Min = min(avg_customer_hrs_out, na.rm = TRUE),
      `25% Quantile` = quantile(avg_customer_hrs_out, probs = 0.25, na.rm = TRUE),
      Median = quantile(avg_customer_hrs_out, probs = 0.5, na.rm = TRUE),
      `75% Quantile` = quantile(avg_customer_hrs_out, probs = 0.75, na.rm = TRUE),
      Max = max(avg_customer_hrs_out, na.rm = TRUE),
      Distribution = list(avg_customer_hrs_out)
    ) %>%
    mutate(across(where(is.numeric), ~ round(.x, 3)))
})
    
bind_rows(hur_info$out_table) %>%
  gt() %>%
  gtExtras::gt_plt_dist(Distribution) %>%
  tab_header(
    title = "Estimated Average Customer Hours without Electricity",
  )
```