---
title: "Large Hurricane Analysis - Outage"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{top_15}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

## Program Background

-   Script name: Large Hurricane Analysis - Outage

-   Project name: ADRD and Hurricane Exposure

-   Program description:

    Outage measure calculation for 15 previously selected large hurricanes. Only hurricanes in 2014 and after had available outage data.

    Number of customers without electricity is collected by EAGLE-I. County-level electric customer estimates are generating by adding the census business survey and ACS5 population estimates for a county and normalizing to state-wide energy customer estimates from the Energy Information Administration.

-   Data:

    -   [EAGLE-I outage data](https://figshare.com/ndownloader/articles/24237376/versions/2)

    -   [Census Business Survey](https://www2.census.gov/programs-surveys/susb/tables/)

    -   ACS 5 - through `tidycensus` package

    -   [EIA State Customer Estimates](https://www.eia.gov/electricity/data/browser/data/index.php?method=getMapData&geosetId=ELEC.CUSTOMERS.ALL.A&freq=A&topic=56)

    -   HURDAT2 - to get storm dates

    -   [2010 ZCTA to County Relationships](https://www2.census.gov/geo/docs/maps-data/data/rel/zcta_county_rel_10.txt)

-   Restrictions:

-   Outline of program/analysis:

    -   Specify all storms

    -   Download and process hurdat2 tracks

    -   Extract outage data for dates specified by the HURDAT2 plus 2 days after

    -   Calculate outage estimates for the maximum percentage of customers without electricity and the average duration of outages

    -   Crosswalk county estimates to ZCTA using 2010 ZCTA to County Relationship file

-   Author: Ryan Zomorrodi

-   Revisions

-   Input files

    -   All files are downloaded using R functions
    -   Cache is stored in `raw-data\eaglei`

-   Output files

    -   `tract/{storm id}/outage.csv`

    -   `zcta/{storm id}/outage.csv`

        | Column | Description |
        |------------------------------------|------------------------------------|
        | `geoid` | 2010 geographic identifier |
        | `max_pct_out` | Maximum percentage of customers out of electricity at any point |
        | `avg_customer_hrs_out` | Average number of hours each customer was out of electricity for |

Start Time

```{r}
start_time <- Sys.time()
cat(start_time, "\n")
```

```{r setup}
library(exposur)
library(tidyverse)
library(tigris)
library(sf)
```

The following hurricanes were selected.

```{r}
hur_info <- readRDS( "raw-data/RF1_Hurricanes.rds")
# get fips codes for each state
fips_codes <- distinct(tigris::fips_codes, state, state_code)
hur_info$fips <- pmap(hur_info, function(states, ...) {
  fips_codes$state_code[match(states, fips_codes$state)]
})
```

## Get hurdat2 info for each storm

```{r}
exposur:::download_newest_hurdat2(path = "raw-data/hurdat2")

hur_info$hurdat2 <- pmap(hur_info, function(hurricane, ...){
  hurdat  <- get_hurdat2(
    storm_id = hurricane,
    path = "raw-data/hurdat2"
  )
  
  list(
    start_date = min(hurdat$datetime),
    end_date = max(hurdat$datetime),
    hurdat = hurdat
  )
})
hur_info <- unnest_wider(hur_info, hurdat2)
```

## Get outage data

```{r}
hur_info$eaglei <- pmap(hur_info, function(start_date, end_date, ...) {
  if (year(start_date) < 2014) {
    return(list(NULL))
  }

  get_eaglei(
    interval = interval(start_date, end_date + days(2)),
    path = "raw-data/eaglei"
  )
})
```

## Calculate outage

```{r}
hur_info$outage <- pmap(hur_info, function(eaglei, fips, start_date, ...) {
  if (year(start_date) < 2014) {
    return(list(NULL))
  }

  calc_outage(eaglei, fips)
})

pwalk(hur_info, function(outage, hurricane, start_date, ...) {
  if (year(start_date) >= 2014) {
    write_csv(outage, str_c("tract/", hurricane, "/outage.csv"))
  }
})
```

## Crosswalk to zctas

```{r}
zcta_to_county <- read_csv("https://www2.census.gov/geo/docs/maps-data/data/rel/zcta_county_rel_10.txt") %>%
  mutate(geoid_z = str_c(STATE, ZCTA5)) %>%
  mutate(geoid_c = str_c(STATE, COUNTY))

outages <- Sys.glob("tract/*/outage.csv") %>%
  map(function(outage_path) {
    outage <- read_csv(outage_path, col_types = cols(geoid = col_character()))
    
    outage %>%
      left_join(zcta_to_county, join_by(geoid == geoid_c)) %>%
      group_by(geoid_z) %>%
      summarise(
        max_pct_out_xwalk = sum(ZPOPPCT * max_pct_out) / 100,
        avg_customer_hrs_out_xwalk = sum(ZPOPPCT * avg_customer_hrs_out) / 100
      )
    
    outage %>%
      write_csv(str_replace(outage_path, "tract", "zcta"))
  })
```

End Time

```{r}
end_time <- Sys.time()
cat("End time: ", end_time, "\n")
cat("Duration: ", start_time - end_time, "\n")
```