---
title: "Large Hurricane Analysis - Precipitation"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{top_15}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Program Background

-   Script name: Large Hurricane Analysis - Precipitation

-   Project name: ADRD and Hurricane Exposure

-   Program description:

    Amount of precipitation was calculated for every day specified within the HURDAT2 plus an additional 2 days after. nClimGrid raster data was used to calculate an areal mean precipitation for all tracts/zctas.

-   Data:

    -   HURDAT2 - to get storm dates
    -   [nClimGrid-Daily](https://www.ncei.noaa.gov/products/land-based-station/nclimgrid-daily)

-   Restrictions:

-   Outline of program/analysis:

    -   Specify all storms
    -   Download and process hurdat2 tracks
    -   Download nClimGrid data for duration of storm plus 2 days after
    -   Calculate areal mean precipitation

-   Author: Ryan Zomorrodi

-   Revisions

-   Input files

    -   All files are downloaded using R functions
    -   Cache is stored in `raw-data\nClimGrid`

-   Output files

    -   `tract/{storm id}/daily_prcp.csv`

    -   `zcta/{storm id}/daily_prcp.csv`

        +-------------+-----------------------------------------------------------------------------------------------------------+
        | Column      | Description                                                                                               |
        +=============+===========================================================================================================+
        | `geoid`     | 2010 geographic identifier                                                                                |
        +-------------+-----------------------------------------------------------------------------------------------------------+
        | `date`      | Date in UTC                                                                                               |
        +-------------+-----------------------------------------------------------------------------------------------------------+
        | `mean_prcp` | Mean areal precipitation (mm) for the specified `geoid` and `tract`                                       |
        |             |                                                                                                           |
        |             | Note: there may be NAs for a small number of coastline tract which were not entirely covered by nClimGrid |
        +-------------+-----------------------------------------------------------------------------------------------------------+

```{r setup, eval = FALSE}
library(exposur)
library(tidyverse)
library(tigris)
library(sf)
```

The following hurricanes were selected.

```{r, eval = FALSE}
hur_info <- readRDS( "raw-data/RF1_Hurricanes.rds")
# get fips codes for each state
fips_codes <- distinct(tigris::fips_codes, state, state_code)
hur_info$fips <- pmap(hur_info, function(states, ...) {
  fips_codes$state_code[match(states, fips_codes$state)]
})
```

## Get hurdat2 info for each storm

```{r, eval = FALSE}
exposur:::download_newest_hurdat2(path = "raw-data/hurdat2")

hur_info$hurdat2 <- pmap(hur_info, function(hurricane, ...){
  hurdat  <- get_hurdat2(
    storm_id = hurricane,
    path = "raw-data/hurdat2"
  )
  
  list(
    start_date = min(hurdat$datetime),
    end_date = max(hurdat$datetime),
    hurdat = hurdat
  )
})
hur_info <- unnest_wider(hur_info, hurdat2)
```

## Get states

```{r, warning = FALSE, eval = FALSE}
hur_info$tracts_sf <- pmap(hur_info, function(start_date, fips, ...) {
  tracts_list <- map(fips, 
    ~ tracts(state = .x, year = 2010, cb = TRUE)
  )
  bind_rows(tracts_list) %>%
    mutate(GEOID = str_sub(GEO_ID, -11)) %>% # get tract geoids
    filter(!st_is_empty(.)) %>% # remove tracts with empty geographies
    st_transform(4326) %>% # transform to WGS 84
    rename(geoid = GEOID)
})

hur_info$zctas_sf <- pmap(hur_info, function(start_date, fips, ...) {
  zctas_list <- map(fips, 
    ~ zctas(state = .x, year = 2010, cb = FALSE)
  )
  bind_rows(zctas_list) %>%
    filter(!st_is_empty(.)) %>% # remove zctas with empty geographies
    st_transform(4326) %>%
    rename(geoid = GEOID10)
})
```

## Get nClimGrid

```{r, eval = FALSE}
hur_info$nClimGrid <- pmap(hur_info, function(start_date, end_date, ...) {
  get_nClimGrid(
    interval = interval(start_date, end_date + days(2)),
    path = "raw-data/nClimGrid"
  )["prcp"]
})
```

## Calculate daily precipitation

```{r, eval = FALSE}
hur_info$tract_prcp_daily_avg <- pmap(hur_info, function(nClimGrid, tracts_sf, ...) {
  calc_daily_avgs(
    nClimGrid = nClimGrid,
    geography = tracts_sf,
    geoid = "geoid"
  )
})

pwalk(hur_info, function(tract_prcp_daily_avg, hurricane, ...) {
  write_csv(tract_prcp_daily_avg, str_c("tract/", hurricane, "/daily_prcp.csv"))
})

hur_info$zcta_prcp_daily_avg <- pmap(hur_info, function(nClimGrid, zctas_sf, ...) {
  calc_daily_avgs(
    nClimGrid = nClimGrid,
    geography = zctas_sf,
    geoid = "geoid"
  )
})

pwalk(hur_info, function(zcta_prcp_daily_avg, hurricane, ...) {
  write_csv(zcta_prcp_daily_avg, str_c("zcta/", hurricane, "/daily_prcp.csv"))
})
```

End Time

```{r, eval = FALSE }
end_time <- Sys.time()
cat("End time: ", end_time, "\n")
cat("Duration: ", start_time - end_time, "\n")
```

## Visualizations

```{r, include = FALSE}
library(tidyverse)
library(tigris)
library(sf)

# repeat from above
hur_info <- readRDS( "../raw-data/RF1_Hurricanes.rds")
fips_codes <- distinct(tigris::fips_codes, state, state_code)
hur_info$fips <- pmap(hur_info, function(states, ...) {
  fips_codes$state_code[match(states, fips_codes$state)]
})

hur_info$prcp <- pmap(hur_info, function(hurricane, ...) {
  read_csv(
    str_c("../tract/", hurricane, "/daily_prcp.csv"),
    col_types = cols(geoid = col_character())
  )
})

hur_info$tract_sf <- pmap(hur_info, function(fips, ...) {
  tract_list <- map(fips, 
    ~ tracts(state = .x, year = 2010, cb = TRUE)
  )
  bind_rows(tract_list) %>%
    mutate(GEOID = str_sub(GEO_ID, -11)) %>%
    filter(!st_is_empty(.))
})

hur_info$states_sf <- pmap(hur_info, function(fips, ...) {
  states(year = 2010, cb = TRUE) %>%
    filter(STATE %in% fips)
})

hur_info$track_line <- pmap(hur_info, function(hurricane, ...) {
  st_read(str_c("../tract/", hurricane, "/track_line.geojson"))
})

hur_info$distance_wind <- pmap(hur_info, function(hurricane, ...) {
  read_csv(
    str_c("../tract/", hurricane, "/distance_wind.csv"),
    col_types = cols(geoid = col_character())
  )})
```

```{r, fig.width = 10, fig.height = 10, echo = FALSE, warning = FALSE}
library(patchwork)

hur_info$prcp_plots <- pmap(hur_info, function(hurricane, tract_sf, distance_wind, prcp, states_sf, track_line,  ...) {
  prcp_sf <- left_join(
    prcp, 
    distance_wind, 
    by = join_by(geoid)
  ) %>%
    mutate(closest_date = as_date(closest_datetime)) %>%
    group_by(geoid) %>%
    filter(date %within% interval(closest_date - day(2), closest_date + day(1))) %>%
    summarize(mean_prcp = sum(mean_prcp)) %>%
    left_join(
      x = tract_sf,
      by = join_by(GEOID == geoid)
    ) %>%
    mutate(quantile_mean_prcp = classInt::classify_intervals(round(mean_prcp, 3), n = 5, style = "quantile"))

  
  ggplot(prcp_sf) +
    geom_sf(aes(fill = quantile_mean_prcp), color = NA) +
    geom_sf(data = states_sf, fill = NA) +
    geom_sf(
      data = st_intersection(
        st_transform(track_line, st_crs(states_sf)), 
        st_as_sfc(st_bbox(states_sf))
      )
    ) +
    scale_fill_brewer(palette = "Blues", na.value= "grey") +
    coord_sf(crs = "ESRI:102003") +
    theme_bw() +
    theme(
      axis.text.x = element_blank(),
      axis.text.y = element_blank(),
      axis.ticks = element_blank()
    ) +
    labs(title = hurricane, fill = "")
})

wrap_plots(
  hur_info$prcp_plots,
  ncol = 3
) +
   plot_annotation(
    title = "Estimated Precipitation (mm)",
    subtitle = "Using Anderson Definitions (2 days before and 1 day after nearest approach)",
    theme = theme(plot.title = element_text(size = 16))
  )
```


```{r, echo = FALSE, warning = FALSE}
library(gt)

hur_info$prcp_table <- pmap(hur_info, function(hurricane, tract_sf, distance_wind, prcp, states_sf, track_line,  ...) {
  left_join(
    prcp, 
    distance_wind, 
    by = join_by(geoid)
  ) %>%
    mutate(closest_date = as_date(closest_datetime)) %>%
    group_by(geoid) %>%
    filter(date %within% interval(closest_date - day(2), closest_date + day(1))) %>%
    summarize(mean_prcp = sum(mean_prcp)) %>%
    ungroup() %>%
    summarize(
      Storm = hurricane,
      Mean = mean(mean_prcp, na.rm = TRUE),
      SD = sd(mean_prcp, na.rm = TRUE),
      Min = min(mean_prcp, na.rm = TRUE),
      `25% Quantile` = quantile(mean_prcp, probs = 0.25, na.rm = TRUE),
      Median = quantile(mean_prcp, probs = 0.5, na.rm = TRUE),
      `75% Quantile` = quantile(mean_prcp, probs = 0.75, na.rm = TRUE),
      Max = max(mean_prcp, na.rm = TRUE),
      `Proportion 75+ mm` = mean(mean_prcp >= 75, na.rm = TRUE),
      Distribution = list(mean_prcp)
    ) %>%
    mutate(across(where(is.numeric), ~ round(.x, 3)))
})
    
bind_rows(hur_info$prcp_table) %>%
  gt() %>%
  gtExtras::gt_plt_dist(Distribution) %>%
  tab_header(
    title = "Estimated Precipitation (mm)",
    subtitle = "Using Anderson Definitions (2 days before and 1 day after nearest approach)"
  )
```