---
title: "Large Hurricane Analysis - Distance and Wind"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{top_15}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

Start Time

```{r}
start_time <- Sys.time()
cat(start_time, "\n")
```

## Program Background

-   Script name: Large Hurricane Analysis - Distance and Wind

-   Project name: ADRD and Hurricane Exposure

-   Program description:

    Maximum wind was calculated for each tract using the `stormwindmodel` published by Anderson at the geocentroid for each tract.

-   Data:

    -   HURDAT2

-   Restrictions:

-   Outline of program/analysis:

    -   Specify all storms
    -   Download and process hurdat2 tracks
    -   Calculate geo-centroid for each tract
    -   Utilize stormwindmodel to produce wind exposure estimates
    -   Output imputed track geojsons for visualization

-   Author: Ryan Zomorrodi

-   Revisions

-   Input files

    -   All files are downloaded using R functions
    -   Cache is stored in `raw-data\hurdat2`

-   Output files

    -   `tract/{storm_id}/distance_wind.csv`

        | Col1 | Col2 |
        |------------------------------------|------------------------------------|
        | geoid | Geographic identifier |
        | closest_datetime | Date and time in UTC that storm is closest to the imputed storm track |
        | closest_dist | Closest distance (km) to imputed central track |
        | highwind_datetime | Date and time in UTC of maximum wind speed |
        | vmax_sust | Maximum 10-m 1-min sustained wind for the tropical cyclone (m/s) |
        | vmax_gust | Max 10-m 1-min gust wind experienced at grid point (m/s) |
        | sust_dur | Duration of time a certain sustained wind was experienced at grid point (minutes) |
        | gust_dur | Duration of time a certain gust wind was experienced at grid point (minutes) |

    -   `tract/{storm_id}/track_line.geojson`

    -   `tract/{storm_id}/track_points.geojson`

```{r setup}
library(exposur)
library(tidyverse)
library(tigris)
library(sf)
```

The following hurricanes were selected.

```{r}
# manually selected states of interest for each storm
hur_info <- tribble(
  ~ hurricane, ~ states,
  "KATRINA-2005",  c("LA", "MS", "AL", "FL", "TN"),
  "HARVEY-2017",   c("TX", "LA"),
  "SANDY-2012",    c("NJ", "DE", "MD", "NY", "CT", "RI", "VA", "MA", "PA"),
  "IRMA-2017",     c("FL", "GA", "SC", "AL"),
  "IKE-2008",      c("TX", "LA"),
  "IVAN-2004",     c("AL", "MS", "FL", "GA", "TN", "NC", "SC"),
  "RITA-2005",     c("TX", "LA"),
  "FLORENCE-2018", c("NC", "SC", "GA", "VA"),
  "MICHAEL-2018",  c("FL", "GA", "AL"),
  "WILMA-2005",    c("FL"),
  "CHARLEY-2004",  c("FL", "GA", "SC", "NC"),
  "IRENE-2011",    c("SC", "NC", "VA", "MD", "DE", "NJ", "PA", "NY", "CT"),
  "FRANCES-2004",  c("FL", "GA", "SC", "NC", "AL", "TN", "VA"),
  "ALLISON-2001",  c("TX", "LA", "MS", "AL", "FL"),
  "MATTHEW-2016",  c("FL", "GA", "SC", "NC", "VA")
)
# get fips codes for each state
fips_codes <- distinct(tigris::fips_codes, state, state_code)
hur_info$fips <- pmap(hur_info, function(states, ...) {
  fips_codes$state_code[match(states, fips_codes$state)]
})
```

## Get hurdat2 info for each storm

```{r}
exposur:::download_newest_hurdat2(path = "raw-data/hurdat2")

hur_info$hurdat2 <- pmap(hur_info, function(hurricane, ...){
  hurdat  <- get_hurdat2(
    storm_id = hurricane,
    path = "raw-data/hurdat2"
  )
  
  list(
    start_date = min(hurdat$datetime),
    end_date = max(hurdat$datetime),
    hurdat = hurdat
  )
})
hur_info <- unnest_wider(hur_info, hurdat2)
```

## Get states

```{r, warning = FALSE}
hur_info$states_sf <- pmap(hur_info, function(start_date, fips, ...) {
  states_list <- map(fips, 
    ~ tracts(state = .x, year = 2010, cb = TRUE)
  )
  bind_rows(states_list) %>%
    mutate(GEOID = str_sub(GEO_ID, -11)) %>% # get tract geoids
    filter(!st_is_empty(.)) %>% # remove tracts with empty geographies
    st_transform(4326) %>% # transform to WGS 84
    st_centroid() # get geo-centroid of each tract
})
```

## Calculate interpolated track

```{r}
hur_info$track <- pmap(hur_info, function(hurdat, ...) {
  calc_interpolated_track(hurdat)
})

pwalk(hur_info, function(track, hurricane, ...) {
  st_write(track, str_c("tract/", hurricane, "/track_points.geojson"))
  
  track %>%
    st_combine() %>%
    st_cast("LINESTRING") %>%
    st_write(str_c("tract/", hurricane, "/track_line.geojson"))
})
```

## Calculate closest distance

```{r}
hur_info$closest_dist <- pmap(hur_info, function(track, states_sf, ...) {
  calc_closest_dist(track, states_sf)
})
```

## Calculate wind speed

```{r}
hur_info$anderson_wind <- pmap(hur_info, function(hurdat, states_sf, ...) {
  calc_anderson_wind(hurdat, states_sf)
})
```

## Join output and write to file

```{r}
pwalk(hur_info, function(closest_dist, anderson_wind, hurricane, ...) {
  full_join(closest_dist, anderson_wind, join_by(geoid)) %>%
    relocate(closest_date, .before = closest_dist) %>%
    rename(highwind_datetime = datetime, closest_datetime = closest_date) %>%
    write_csv(str_c("tract/", hurricane, "/distance_wind.csv"))
})
```

End Time

```{r}
end_time <- Sys.time()
cat("End time: ", end_time, "\n")
cat("Duration: ", start_time - end_time, "\n")
```